---
title: "Create column based on datecolumn values in R"
author: "Naren Srinivasan"
date: "9 December 2018"
output: html_document
---
# The problem

\link{https://stackoverflow.com/questions/53689628/create-column-based-on-datecolumn-values-in-r}

I wrote a function that creates column based on a datetime column using parameters starting and ending dates, but I can't get it to work.

df is a data frame object.

```{r eval = F}
create_gv<-function(df, s_ymd, e_ymd, char) {
  df<-get(df)
  for (i in (1:nrow(df))) {
  ymd <- format(df[i,1],"%y%m%d") 
  if ((strptime(ymd,format = "%y%m%d") >= strptime(s_ymd,format = "%y%m%d") & strptime(ymd,format = "%y%m%d") <= strptime(e_ymd,format = "%y%m%d")) == TRUE) {
    df$group_var[i]<-char
  }
  }
}

create_gv("example","171224","171224","D")
```

I get:

`
> example
           start_time group_var
1 2017-12-24 10:42:39        NA
2 2017-12-24 10:44:31        NA
3 2018-01-14 12:05:53        NA
4 2018-01-14 12:22:12        NA
`

Reproducible data frame named example here:

```{r eval = F}
structure(list(start_time = structure(c(1514112159, 1514112271, 
1515931553, 1515932532), class = c("POSIXct", "POSIXt"), tzone = ""), 
    group_var = c(NA, NA, NA, NA)), .Names = c("start_time", 
"group_var"), row.names = c(NA, -4L), class = "data.frame")
```


Wanted output:

           start_time group_var
1 2017-12-24 10:42:39         D
2 2017-12-24 10:44:31         D
3 2018-01-14 12:05:53         NA
4 2018-01-14 12:22:12         NA

# The solution
```{r}

df <- structure(list(start_time = structure(c(1514112159, 1514112271, 
1515931553, 1515932532), class = c("POSIXct", "POSIXt"), tzone = ""), 
    group_var = c(NA, NA, NA, NA)), .Names = c("start_time", 
"group_var"), row.names = c(NA, -4L), class = "data.frame")
df
```

From your description, my understanding is that you want to check if the date in a row  is between the start and end date (which are scalars), and update the value of `group_var` accordingly.

The `lubridate` package provides a set of tools which allow to easily work with dates. In order to compare dates you don't need to format them. `format` only helps with the viewing of these dates. I have used to the `dplyr` package which allows you to easily perform data transformations.

To solve the problem we use the `dplyr::mutate` function which transforms a column by row, as a function of other columns. In this case, tthe date column in our dataset is to compared with scalar start and end times in order to codify the `group_var` variable.

```{r}
library(lubridate)

char <- "D"
# Randomly setting the start and end times for the purpose of the example. Any value can be passed to this.
s_ymd <- df$start_time[1] - 5000
e_ymd <- df$start_time[2] + 5000


df %>% dplyr::mutate(group_var = ifelse(start_time > s_ymd & start_time < e_ymd,
                                        char, NA)) -> df
df
```

To use a function write:

```{r}
create_gv <- function(start_time, s_ymd, e_ymd, char){
  g_var <- ifelse(start_time > s_ymd & start_time < e_ymd,
                                        char, NA)
  return(g_var)
}

df %>% dplyr::mutate(group_var = create_gv(start_time, !!s_ymd, !!e_ymd, !!char)) 
```

sHere since s_ymd, e_ymd and char are scalars (i.e. not columns in the data frame), we need to unquote them

Note - The `%>%` operator is used to make the code more readable by piping sequence of operations one after another, and comes from the `magrittr` package. This is of course not required
