---
title: "Select variables from list of tibbles using names in each tibble"
author: "Naren Srinivasan"
date: "25 November 2018"
output: html_document
---

# The problem

\link{https://stackoverflow.com/questions/53464733/select-variables-from-list-of-tibbles-using-names-in-each-tibble}


I have a list of tibbles or data frames (each one with a name in the list) and I want to:

* create new columns with the same name long.col in each tibble from variables with different names
* do this by matching from another tibble with the key linking tibble name and column name in each tibble and finally
* binding rows for all the newly created columns with the same name long.col in each tibble as well as identifying the original tibble where they came from.

I would like to do this preferably using tidyverse functions. Here is a an example of both: 

* the list of tibbles
* the key tibble identifying tibble names and variables to select in each one

```{r}
library(dplyr)
df1 <- tibble(v1 = c(rep("A", 5), rep("B", 5)),
          v2 = 1:10)
df2 <- tibble(v1 = c(rep("C", 6), rep("D", 6)),
          v3 = 11:22)
df3 <- tibble(v1 = c(rep("E", 4), rep("F", 4)),
          v4 = 23:30)

list.df <- list(df1, df2, df3)
names(list.df) <- c("data1", "data2", "data3")

key <- tibble(data = c("data1", "data2", "data3"),
          vars = c("v2", "v3", "v4"))
```

The final output should look like:

```{r}
 final.df <- tibble(data = c(rep("data1", 10), rep("data2", 12), rep("data3", 8)),
               long.col = 1:30)
```

# The solution

```{r}
getBaseVar <- function(dfName, keyTibble){
  varToBeTransformed <- keyTibble %>% dplyr::filter(data == (!!dfName)) %>% 
                           dplyr::select(vars) %>% dplyr::first() %>% rlang::sym(.)
  return(varToBeTransformed)

}

list.df.transformed <- purrr::imap(list.df, function(df, name){
  df %>% dplyr::mutate( long.col := !!getBaseVar(name, key))
})

final.df <- tibble(data = character(), long.col = numeric()) 
purrr::imap(list.df.transformed, function(df, name){
  repeatedNameCol <- tibble(data = rep(name, nrow(df)))
  dataToBind <- df %>% dplyr::select(long.col) %>% dplyr::bind_cols(repeatedNameCol)
  return(dataToBind)
}) %>% dplyr::bind_rows(.) -> final.df
final.df
```


# Top-voted solution

```{r}
out <- purrr::pmap_df(list(list.df,
                    names(list.df),
                    key[['vars']]), ~ tibble(data = ..2, long.col = ..1[[..3]]))
out
```

